# Activity monitoring
rucrbser

## Setting directory

```{r directory}
setwd("C:/users/x/desktop/PA1_template")
```


## Setting global option

```{r setoption, echo = TRUE}
library(knitr)
opts_chunk$set(echo = TRUE)
```


## Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?

```{r}
stepstotal <- with(activity, tapply(steps, date, sum))
stepstotal <- stepstotal[!is.na(stepstotal)]

hist(stepstotal)

stepsmean <- as.character(round(mean(stepstotal), digits = 2))
stepsmedian <- median(stepstotal)
```

The mean and median of the total number of steps 
taken per day is `r stepsmean` and `r stepsmedian`.


## What is the average daily activity pattern?

```{r}
activity <- transform(activity, interval = factor(interval))
stepsaverage <- with(activity, tapply(steps, interval, mean, na.rm = TRUE))
with(activity, plot(levels(interval), stepsaverage, type = "l",
                    xlab = "5-minute interval", ylab = "average steps"))

intervalmax <- levels(activity$interval)[which.max(stepsaverage)]
```

The `r intervalmax` 5-minute interval contains the maximum number of steps 
on average across all the days in the dataset.


## Imputing missing values

```{r}
bad <- !(complete.cases(activity))
NAnum <- sum(bad)

library(dplyr)
activity2 <- as.data.frame(group_by(activity, interval))
activity2[bad, ]$steps <- stepsaverage

stepstotal2 <- with(activity2, tapply(steps, date, sum))
hist(stepstotal2)
stepsmean2 <- as.character(round(mean(stepstotal2), digits = 2))
stepsmedian2 <- as.character(round(median(stepstotal2), digits = 2))
```

The total number of missing values in the dataset is `r NAnum`.

The mean and median of the total number of steps taken per day is `r stepsmean2` and `r stepsmedian2`. They are much close to the estimates from the first part of the assignment. There is no much impact imputing missing data on the estimates of the total daily number of steps with the mean/median for that day or the mean for that 5-minute interval.


## Are there differences in activity patterns between weekdays and weekends?

```{r fig.height=10}
Sys.setlocale("LC_ALL", "English")
activity2 <- mutate(activity2, day = weekdays(as.Date(activity2$date)))
activity2$day[!((activity2$day %in% 'Saturday')|
                      (activity2$day %in% 'Sunday'))] = "weekday"
activity2$day[(activity2$day %in% 'Saturday')|
                      (activity2$day %in% 'Sunday')] = "weekend"
activity2 <- transform(activity2, day = factor(day))

library(lattice)
stepsaveweekday <- with(activity2[activity2$day == "weekday", ], 
                        tapply(steps, interval, mean))
stepsaveweekend <- with(activity2[activity2$day == "weekend", ], 
                        tapply(steps, interval, mean))
par(mfrow = c(2,1))
with(activity2, plot(levels(interval), stepsaveweekday, type = "l", col = "blue",
                     xlab = "interval", ylab = "average steps", main = "weekday"))
with(activity2, plot(levels(interval), stepsaveweekend, type = "l", col = "blue",
                     xlab = "interval", ylab = "average steps", main = "weekend"))
```

